---
title: "Phase 3 Preliminary Report"
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(DT)
```

```{r data}
x <- read_csv("phase3_airtable.csv") %>% 
   rename(name = `Name (Last, First)`, 
         degree = `HLS Degree(s)`,
         year = `Grad Date(s)`,
         category = `Position / Role`, 
         race = `Race`, 
         gender = `Gender`, 
         party = `Party`, 
         country = `Country`,
         priority = `Priority Level`) 
```

# Priority Levels

The priority levels divide the entries pretty evenly.

```{r levels}
x %>% 
  filter(is.na(priority) == FALSE) %>% 
  group_by(priority) %>% 
  mutate(priority = as.numeric(str_replace_all(priority, "[^1-9]", ""))) %>% 
  summarize(priority_total = n()) %>% 
  rename(`Priority Level` = priority,
         Total = priority_total) -> a

datatable(a, rownames = FALSE)
```

# Gender in Priority Levels

As priority level decreases, the ratio of males to females increases. This is 
expected based on the analysis after Phase 2: in Public Interest, the 
main role in Priority 1, women far outnumbered men; men were better represented in
Government and Corporate and formed a small majority in Judiciary and Law Firm.

```{r gender}
x %>% 
  filter(priority == 1) %>% 
  ggplot(aes(x = gender)) +
  geom_bar() +
  labs(title = "Priority Level 1: Gender", 
       fill = "Gender",
       x = "", 
       y = "")

x %>% 
  filter(priority == 2) %>% 
  ggplot(aes(x = gender)) +
  geom_bar() +
  labs(title = "Priority Level 2: Gender", 
       fill = "Gender",
       x = "", 
       y = "")

x %>% 
  filter(priority == 3) %>% 
  ggplot(aes(x = gender)) +
  geom_bar() +
  labs(title = "Priority Level 3: Gender", 
       fill = "Gender",
       x = "", 
       y = "")
```

# Roles in Priority Levels

Note that though anyone who has Public Interest role was included 
in Priority 1, such a person, for example, may also have a Corporate role which
would also appear in the Priority 1 chart. 

Note also that some exceptions were made to the priority criteria. The reasons 
for these are given in parentheses on the AirTable spreadsheet.

```{r role}
x %>% 
  separate_rows(category, sep = ",") %>% 
  mutate(category = str_replace_all(category, "Public Interest Law / Nonprofit Legal Services", "Public Interest")) %>% 
  filter(priority == 1) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(category)))) +
  scale_x_continuous(breaks = seq(0, 70, 10)) +
  geom_bar() +
  labs(title = "Priority Level 1: Roles", 
       x = "", 
       y = "")

x %>% 
  separate_rows(category, sep = ",") %>% 
  filter(priority == 2) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(category)))) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 70, 10)) +
  labs(title = "Priority Level 2: Roles", 
       x = "", 
       y = "")

x %>% 
  separate_rows(category, sep = ",") %>% 
  filter(priority == 3) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(category)))) +
  geom_bar() +
  scale_x_continuous(breaks = seq(0, 70, 10)) +
  labs(title = "Priority Level 3: Roles", 
       x = "", 
       y = "")
```

# Race in Priority Levels

The racial composition of Priority Level 1 is roughly equal to the racial composition 
in all the entries, with Blacks making up a slightly smaller proportion compared to all entries
and other minorities making up slighty more. 

In Priority Levels 2 and 3, blacks form a plurality, mostly due to the fact that these
levels contain the CBA Judges list.

```{r race}
x %>% 
  filter(priority == 1) %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian",
                          str_detect(race, "White") == TRUE ~ "White")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(race)), fill = race)) +
  geom_bar(aes(x = after_stat(count/sum(count)))) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, .5, .05)) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White", "American Indian", "Aboriginal"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat", "grey", "grey"),
                    na.value = "grey") +
  labs(title = "Priority Level 1: Race", 
       x = "", 
       y = "",
       fill = "Race")

x %>% 
  filter(priority == 2) %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian",
                          str_detect(race, "White") == TRUE ~ "White")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(race)), fill = race)) +
  geom_bar(aes(x = after_stat(count/sum(count)))) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, .5, .05)) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat"),
                    na.value = "grey") +
  labs(title = "Priority Level 2: Race", 
       x = "", 
       y = "",
       fill = "Race")

x %>% 
  filter(priority == 3) %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian",
                          str_detect(race, "White") == TRUE ~ "White")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(race)), fill = race)) +
  geom_bar(aes(x = after_stat(count/sum(count)))) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, .5, .05)) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat"),
                    na.value = "grey") +
  labs(title = "Priority Level 3: Race", 
       x = "", 
       y = "",
       fill = "Race")
```

# Decade in Priority Levels

Priority Level 1 is the decade with the greatest percentage of degrees earned 
in the '90s or after. Priority Level 2 has the least of these, and the highest
percentage of degrees earned in the '70s. 

Note that the few degrees earned before 1950 have been excluded from all charts for easier
comparison.

```{r decade}
x %>% 
  filter(priority == 1) %>% 
  separate_rows(year, sep = "\\s*/\\s*") %>% 
  mutate(year = as.numeric(str_replace_all(year, "[^0-9].*", ""))) %>% 
  filter(year >= 1950) %>% 
  ggplot(aes(y = year)) +
  geom_histogram(aes(x = after_stat(count/sum(count))),
                 boundary = 1950, binwidth = 10, color = "white") +
  scale_y_continuous(breaks = seq(1875, 2015, 10),
                     labels = c("1870-1879", "1880-1889", "1890-1899", "1900-1909",
                                "1910-1919","1920-1929", "1930-1939", "1940-1949",
                                "1950-1959","1960-1969", "1970-1979", "1980-1989",
                                "1990-1999", "2000-2009", "2010-2019")) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, .5, .05)) +
    labs(title = "Priority Level 1: Decade", 
       x = "", 
       y = "")

x %>% 
  filter(priority == 2) %>% 
  separate_rows(year, sep = "\\s*/\\s*") %>% 
  mutate(year = as.numeric(str_replace_all(year, "[^0-9].*", ""))) %>% 
  filter(year >= 1950) %>% 
  ggplot(aes(y = year)) +
  geom_histogram(aes(x = after_stat(count/sum(count))),
                 boundary = 1950, binwidth = 10, color = "white") +
  scale_y_continuous(breaks = seq(1875, 2015, 10),
                     labels = c("1870-1879", "1880-1889", "1890-1899", "1900-1909",
                                "1910-1919","1920-1929", "1930-1939", "1940-1949",
                                "1950-1959","1960-1969", "1970-1979", "1980-1989",
                                "1990-1999", "2000-2009", "2010-2019")) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, .5, .05)) +
    labs(title = "Priority Level 2: Decade", 
       x = "", 
       y = "")


x %>% 
  filter(priority == 3) %>% 
  separate_rows(year, sep = "\\s*/\\s*") %>% 
  mutate(year = as.numeric(str_replace_all(year, "[^0-9].*", ""))) %>% 
  filter(year >= 1950) %>% 
  ggplot(aes(y = year)) +
  geom_histogram(aes(x = after_stat(count/sum(count))),
                 boundary = 1950, binwidth = 10, color = "white") +
  scale_y_continuous(breaks = seq(1875, 2015, 10),
                     labels = c("1870-1879", "1880-1889", "1890-1899", "1900-1909",
                                "1910-1919","1920-1929", "1930-1939", "1940-1949",
                                "1950-1959","1960-1969", "1970-1979", "1980-1989",
                                "1990-1999", "2000-2009", "2010-2019")) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, .5, .05)) +
    labs(title = "Priority Level 3: Decade", 
       x = "", 
       y = "")
```

# Degrees in Priority Levels

As intended, Priority Level 1 has the greatest diversity of degrees. One notable
exception is for J.D./M.B.A, the vast majority of whom are in Priority Level 3, largely
because most of them have either Corporate/Business or Law Firm roles. 

```{r degree}
x %>% 
  filter(priority == 1) %>% 
  mutate(degree = str_replace_all(degree, ",", " & ")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(degree)))) +
  geom_bar(aes(x = after_stat(count/sum(count)))) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 1, .1)) +
      labs(title = "Priority Level 1: Degrees", 
       x = "", 
       y = "")

x %>% 
  filter(priority == 2) %>% 
  mutate(degree = str_replace_all(degree, ",", " & ")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(degree)))) +
  geom_bar(aes(x = after_stat(count/sum(count)))) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 1, .1)) +
      labs(title = "Priority Level 2: Degrees", 
       x = "", 
       y = "")

x %>% 
  filter(priority == 3) %>% 
  mutate(degree = str_replace_all(degree, ",", " & ")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(degree)))) +
  geom_bar(aes(x = after_stat(count/sum(count)))) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 1, .1)) +
      labs(title = "Priority Level 3: Degrees", 
       x = "", 
       y = "")
```



