---
title: "Phase 1 Report"
site: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

```{r}
x <- read_csv("phase1_airtable.csv", col_types = cols()) %>% 
  rename(name = `Name (Last, First)`, 
         degree = `HLS Degree(s)`,
         year = `Grad Date(s)`,
         category = `Position / Role`, 
         race = `Race`, 
         gender = `Gender`, 
         party = `Party`, 
         country = `Country`) %>% 
  select(name, degree, year, category, race, gender, party, country) %>% 
  drop_na(name)
```

# Gender

89 entries are Male, 52 are Female, and 1, Jack F. Trope, is likely
male but this has not been included because of the lack of a source. 

```{r gender-hist}
x %>% 
  ggplot(aes(gender)) +
  geom_bar(aes(y = after_stat(count / sum(count)))) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Gender", 
       x = "",
       y = "Number") 
```

# Decades

As expected, the vast majority of entries are from recent decades. The '70s and
'80s have the most degrees. The '50s and '60s are comparatively lacking. It 
may be natural that the most recent graduates are also lacking. 

```{r decade-hist, warning = FALSE}
x %>% 
  separate_rows(year, sep = "\\s*/\\s*") %>% 
  mutate(year = str_replace_all(year, "[^0-9].*", "")) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(y = year)) +
  geom_histogram(boundary = 1950, binwidth = 10, color = "white") +
  scale_y_continuous(breaks = seq(1875, 2015, 10),
                     labels = c("1870-1879", "1880-1889", "1890-1899", "1900-1909",
                                "1910-1919","1920-1929", "1930-1939", "1940-1949",
                                "1950-1959","1960-1969", "1970-1979", "1980-1989",
                                "1990-1999", "2000-2009", "2010-2019")) +
  scale_x_continuous(breaks = seq(0, 40, 5)) +
  labs(title = "Decades",
       x = "Number of Degrees",
       y = "")
```

As expected, gender balance becomes more equal in later decades. It is notable, 
however, that females significantly outnumber males among most recent graduates. 

```{r decade-gender, warning = FALSE}
x %>%
  separate_rows(year, sep = "\\s*/\\s*") %>%
    mutate(year = str_replace_all(year, "[^0-9].*", "")) %>% 
  mutate(year = as.numeric(year)) %>%
  drop_na(year) %>% 
  mutate(decade = case_when(year >= 2010 & year <= 2019 ~ "2010-2019",
                            year >= 2000 & year <= 2009 ~ "2000-2009",
                            year >= 1990 & year <= 1999 ~ "1990-1999",
                            year >= 1980 & year <= 1989 ~ "1980-1989",
                            year >= 1970 & year <= 1979 ~ "1970-1979",
                            year >= 1960 & year <= 1969 ~ "1960-1969",
                            year >= 1950 & year <= 1969 ~ "1950-1959",
                            year >= 1940 & year <= 1949 ~ "1940-1949",
                            year >= 1930 & year <= 1939 ~ "1930-1939",
                            year >= 1920 & year <= 1929 ~ "1920-1929",
                            year >= 1910 & year <= 1919 ~ "1910-1919",
                            year >= 1870 & year <= 1879 ~ "1870-1879")) %>% 
  group_by(decade) %>% 
  mutate(total_in_decade = n()) %>% 
  mutate(percent_of_decade = 1 / total_in_decade) %>% 
  ggplot(aes(y = decade, x = percent_of_decade,  fill = gender, color =)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Decades and Gender", 
       x = "Percent of Gender in Decade", 
       y = "", 
       fill = "Gender", 
       color = "Gender")
  
```

# Races

Among all the entries, blacks outnumber other races. Among only the Micro Museum 
list, whites are the the greatest in number, forming about 40%. 

```{r race-hists}
x %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>%
  ggplot(aes(race)) +
  geom_bar(aes(y = after_stat(count / sum(count)))) +
  labs(title = "Races of All Entries",
       x = "",
       y = "Percent") +
  scale_y_continuous(labels = scales::percent_format())

x %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  slice(1:91) %>% 
  ggplot(aes(race)) +
  geom_bar(aes(y = after_stat(count / sum(count)))) +
  labs(title = "Races of Micro Museum Entries Only",
       x = "",
       y = "Percent") +
  scale_y_continuous(labels = scales::percent_format())
```

Among all races, males outnumber females. The only exceptions are the
races represented by a single member, American Indians and  Australian Aboriginals. 

```{r race-gender}
x %>% 
  drop_na(race) %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  group_by(race) %>% 
  mutate(total_in_race = n()) %>% 
  mutate(percent_of_race = 1 / total_in_race) %>% 
  mutate(test_sum = sum(percent_of_race)) %>% 
  ggplot(aes(x = percent_of_race, y = race, fill = gender, color = gender)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Races and Gender", 
       x = "Percent", 
       y = "",
       fill = "Gender",
       color = "Gender")
```

# Countries

In the Airtable, multiple countries may be listed for one entry. These include 
places where people are involved or born, from what I gathered from a quick
reading of one source, usually Wikipedia. These countries are generally in order
of precedence. The first, and so primary, country for each entry is used here. 

As expected, the vast majority of entries are based in the U.S. Asian countries
make up the the majority of foreign locations. 

```{r country-hist}
x %>% 
  mutate(country = str_replace_all(country, ",.*", "")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(country)))) +
  geom_bar() +
  labs(title = "Countries", 
       x = "Number",
       y = "")
```

# Parties

40 entries have at least one party. If a person has more than one party, each
is counted here. For the most part, I listed parties for those
who had one displayed on Wikipedia, i.e. primarily elected officials. 

Democrats are the most represented party. Several of the parties listed below
are for foreign countries. 

```{r party-hist, warning = FALSE}
x %>% 
  separate_rows(party, sep = "\\s*,\\s*") %>% 
  mutate(party = str_replace_all(party, "Independent", "Independent (Irish)")) %>% 
  drop_na(party) %>% 
  ggplot(aes(y = party)) +
  geom_bar() +
  scale_y_discrete(limits = rev(c("D", "R", "I", "L", "Green", "Democratic Progressive Party",
                              "Independent (Irish)", "Kuomintang", "National Democratic Party of Tibet", 
                              "Popular Democratic", "Reform")),
                   labels = rev(c("Democrat", "Republican", "Independent", "Libretarian", 
                              "Green", "Democratic Progressive Party", "Independent (Irish)",
                              "Kuomintang", "National Democratic Party of Tibet", 
                              "Popular Democratic", "Reform"))) +
  labs(title = "Parties", 
       x = "Number", 
       y = "")
```

Democrats are nearly equal in gender. Compared to the gender ratio among all the entries, women are overrepresented among Democrats and underrepresented among Republicans. 

```{r party-gender}
x %>% 
  separate_rows(party, sep = "\\s*,\\s*") %>% 
  drop_na(party) %>% 
  mutate(party = str_replace_all(party, "Independent", "Independent (Irish)")) %>%
  filter(party %in% c("D", "R", "I", "L")) %>% 
  group_by(party) %>% 
  mutate(total_in_party = n()) %>% 
  mutate(percent_of_party = 1 / total_in_party) %>% 
  ggplot(aes(percent_of_party, party, fill = gender, color = gender)) +
  geom_col() +
  scale_y_discrete(labels = c("Democrat", "Independent", "Libretarian", "Republican")) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Parties and Gender",
       x = "Percent", 
       y = "", 
       fill = "Gender",
       color = "Gender")
```

Both Republicans and Democrats have comparable percentages
of racial minorties, though the particular minorities differ. 

```{r party-race}
x %>% 
   mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  filter(race %in% c("Jewish", "Black", "White", "Hispanic", "Asian")) %>% 
  mutate(race = factor(race,
                       levels = c("Black", "Asian", "Hispanic", "Jewish", "White"))) %>% 
  separate_rows(party, sep = "\\s*,\\s*") %>% 
  drop_na(party) %>% 
  mutate(party = str_replace_all(party, "Independent", "Independent (Irish)")) %>%
  filter(party %in% c("D", "R", "I", "L")) %>% 
  group_by(party) %>% 
  mutate(total_in_party = n()) %>% 
  mutate(percent_of_party = 1 / total_in_party) %>% 
  ggplot(aes(percent_of_party, party, fill = race, color = race)) +
  geom_col() +
  scale_y_discrete(labels = c("Democrat", "Independent", "Libretarian", "Republican")) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  scale_color_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  labs(title = "Parties and Race",
       x = "Percent", 
       y = "",
       fill = "Race",
       color = "Race")
```

# Roles

The roles now are very rough. They have been gathered from a quick reading of 
one source, usually Wikipedia. There are several grey areas: the intersection of Politics,
Public Interest Law, and Activist; most lawyers have at least some experience in a 
firm, but they do not now all have the "Law Firm" role. 

Among all the entries, judges outnumber other roles. Among only the Micro Museum 
list, "politicians" are the the greatest in number. The "Business" 
and "Law Firm" roles are comparatively lacking.  


```{r role-hists}
x %>% 
  separate_rows(category, sep = ",") %>% 
  ggplot(aes(y = fct_rev(fct_infreq(category)))) +
  geom_bar() +
  labs(title = "Roles of All Entries", 
       x = "Number", 
       y = "")

x %>% 
  slice(1:91) %>% 
  separate_rows(category, sep = ",") %>% 
  ggplot(aes(y = fct_rev(fct_infreq(category)))) +
  geom_bar() +
  labs(title = "Roles of Micro Museum Only", 
       x = "Number", 
       y = "")
```

Several roles are around 60% male, which is the proportion of males among all the entries. 

```{r role-gender}
x %>% 
  separate_rows(category, sep = ",") %>% 
  group_by(category) %>% 
  mutate(total_in_category = n()) %>% 
  filter(total_in_category > 1) %>% 
  mutate(percent_of_category = 1 / total_in_category) %>% 
  ggplot(aes(percent_of_category, category, fill = gender, color = gender)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Roles and Gender", 
       x = "Percent", 
       y = "", 
       fill = "Gender",
       color = "Gender")
```

```{r role-race}
x %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  filter(race %in% c("Jewish", "Black", "White", "Hispanic", "Asian")) %>% 
  mutate(race = factor(race,
                       levels = c("Black", "Asian", "Hispanic", "Jewish", "White"))) %>% 
  separate_rows(category, sep = ",") %>% 
  group_by(category) %>% 
  mutate(total_in_category = n()) %>% 
  filter(total_in_category > 1) %>% 
  mutate(percent_of_category = 1 / total_in_category) %>% 
  ggplot(aes(percent_of_category, category, fill = race, color = race)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  scale_color_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  labs(title = "Roles and Race", 
       x = "Percent", 
       y = "", 
       fill = "Race", 
       color = "Race")
```

# Degrees

As expected, the majority of degrees earned are J.Ds. 

```{r degree-hist}
x %>% 
  mutate(degree = str_replace_all(degree, ",", " & ")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(degree)))) +
  geom_bar() +
  labs(title = "Degrees", 
       x = "Number", 
       y = "")
```

There is some but not much overlap between when LL.Bs and J.Ds are earned. No LL.Bs 
are earned after the '60s and a few J.Ds are held by those who graduated even as
early as the '40s.

```{r degree-decade}
x %>% 
   mutate(decade = case_when(year >= 2010 & year <= 2019 ~ "2010-2019",
                            year >= 2000 & year <= 2009 ~ "2000-2009",
                            year >= 1990 & year <= 1999 ~ "1990-1999",
                            year >= 1980 & year <= 1989 ~ "1980-1989",
                            year >= 1970 & year <= 1979 ~ "1970-1979",
                            year >= 1960 & year <= 1969 ~ "1960-1969",
                            year >= 1950 & year <= 1969 ~ "1950-1959",
                            year >= 1940 & year <= 1949 ~ "1940-1949",
                            year >= 1930 & year <= 1939 ~ "1930-1939",
                            year >= 1920 & year <= 1929 ~ "1920-1929",
                            year >= 1910 & year <= 1919 ~ "1910-1919",
                            year >= 1870 & year <= 1879 ~ "1870-1879")) %>% 
  group_by(degree) %>% 
  mutate(total_in_degree = n()) %>% 
  mutate(percent_of_degree = 1 / total_in_degree) %>% 
  filter(degree %in% c("J.D.", "LL.B.", "LL.M.")) %>% 
  ggplot(aes(percent_of_degree, degree, fill = decade, color = decade)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Degrees and Decade", 
       x = "Percent",
       y = "", 
       fill = "Decade",
       color = "Decade")
```