---
title: "Phase 2 Report"
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

```{r}
x <- read_csv("phase1b_airtable.csv", col_types = cols()) %>% 
  rename(name = `Name (Last, First)`, 
         degree = `HLS Degree(s)`,
         year = `Grad Date(s)`,
         category = `Position / Role`, 
         race = `Race`, 
         gender = `Gender`, 
         party = `Party`, 
         country = `Country`) %>% 
  select(name, degree, year, category, race, gender, party, country) %>% 
  drop_na(name)

y <- read_csv("phase1_airtable.csv", col_types = cols()) %>% 
  rename(name = `Name (Last, First)`, 
         degree = `HLS Degree(s)`,
         year = `Grad Date(s)`,
         category = `Position / Role`, 
         race = `Race`, 
         gender = `Gender`, 
         party = `Party`, 
         country = `Country`) %>% 
  select(name, degree, year, category, race, gender, party, country) %>% 
  drop_na(name)
```

# Gender

157 entries are Male, 190 are Female, and 1, Jack F. Trope, is likely
male but this has not been included because of the lack of a source. 

```{r gender-hist}
x %>% 
  ggplot(aes(gender)) +
  geom_bar(aes(y = after_stat(count / sum(count)))) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Gender", 
       x = "",
       y = "Number") 
```

# Decades

Compared to Phase 1, the alumni in the list have bcome more recent. Graduates from
the '80s are now far greater than those from the '70s, and graduates from 
the last two decades now outnumber thos from the '50s and '60s.

```{r decade-hist, warning = FALSE}
x %>% 
  separate_rows(year, sep = "\\s*/\\s*") %>% 
  mutate(year = str_replace_all(year, "[^0-9].*", "")) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(y = year)) +
  geom_histogram(boundary = 1950, binwidth = 10, color = "white") +
  scale_y_continuous(breaks = seq(1875, 2015, 10),
                     labels = c("1870-1879", "1880-1889", "1890-1899", "1900-1909",
                                "1910-1919","1920-1929", "1930-1939", "1940-1949",
                                "1950-1959","1960-1969", "1970-1979", "1980-1989",
                                "1990-1999", "2000-2009", "2010-2019")) +
  labs(title = "Decades Phase 2",
       x = "Number of Degrees",
       y = "")

y %>% 
  separate_rows(year, sep = "\\s*/\\s*") %>% 
  mutate(year = str_replace_all(year, "[^0-9].*", "")) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(y = year)) +
  geom_histogram(boundary = 1950, binwidth = 10, color = "white") +
  scale_y_continuous(breaks = seq(1875, 2015, 10),
                     labels = c("1870-1879", "1880-1889", "1890-1899", "1900-1909",
                                "1910-1919","1920-1929", "1930-1939", "1940-1949",
                                "1950-1959","1960-1969", "1970-1979", "1980-1989",
                                "1990-1999", "2000-2009", "2010-2019")) +
  labs(title = "Decades Phase 1",
       x = "Number of Degrees",
       y = "")
```

In Phase 1, women made up less thn 30% of graduates in the '70s and '80s. Women 
now slightly outnumber men in the '80s and are slightly fewer than men in the '70s.

```{r decade-gender, warning = FALSE}
x %>%
  separate_rows(year, sep = "\\s*/\\s*") %>%
    mutate(year = str_replace_all(year, "[^0-9].*", "")) %>% 
  mutate(year = as.numeric(year)) %>%
  drop_na(year) %>% 
  mutate(decade = case_when(year >= 2010 & year <= 2019 ~ "2010-2019",
                            year >= 2000 & year <= 2009 ~ "2000-2009",
                            year >= 1990 & year <= 1999 ~ "1990-1999",
                            year >= 1980 & year <= 1989 ~ "1980-1989",
                            year >= 1970 & year <= 1979 ~ "1970-1979",
                            year >= 1960 & year <= 1969 ~ "1960-1969",
                            year >= 1950 & year <= 1969 ~ "1950-1959",
                            year >= 1940 & year <= 1949 ~ "1940-1949",
                            year >= 1930 & year <= 1939 ~ "1930-1939",
                            year >= 1920 & year <= 1929 ~ "1920-1929",
                            year >= 1910 & year <= 1919 ~ "1910-1919",
                            year >= 1870 & year <= 1879 ~ "1870-1879")) %>% 
  group_by(decade) %>% 
  mutate(total_in_decade = n()) %>% 
  mutate(percent_of_decade = 1 / total_in_decade) %>% 
  ggplot(aes(y = decade, x = percent_of_decade,  fill = gender, color =)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Decades and Gender", 
       x = "Percent of Gender in Decade", 
       y = "", 
       fill = "Gender", 
       color = "Gender")
```

# Races

Among all the entries, whites are the greatest, forming a plurality slightly over 
30%. Blacks and Hispanics are the most represented minorities, forming slightly 
more and less than 25% respectively. 

Compared to all the entries in Phase 1, the biggest change is that blacks are no
longer near the majority. Also, Asians and not Hispanics were the second most
represented minority, though the percentage of Asians has not changed by much 
between both phases. 

Compared to only those entries from the Micro Museum list, there are more 
minorities in these entires.

```{r race-hists}
x %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian",
                          str_detect(race, "White") == TRUE ~ "White")) %>%
  ggplot(aes(race)) +
  geom_bar(aes(y = after_stat(count / sum(count)))) +
  labs(title = "Races of All Entries",
       x = "",
       y = "Percent") +
  scale_y_continuous(labels = scales::percent_format())

y %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian",
                          str_detect(race, "White") == TRUE ~ "White")) %>% 
  slice(1:91) %>%
  ggplot(aes(race)) +
  geom_bar(aes(y = after_stat(count / sum(count)))) +
  labs(title = "Races of Micro Museum Entries Only",
       x = "",
       y = "Percent") +
  scale_y_continuous(labels = scales::percent_format())
```

Among all races, except Jews and Hispanics, women outnumber men. The opposite
was generally true in the Phase 1. The disparity between men and women is  greatest
among whites. 

```{r race-gender}
x %>% 
  drop_na(race) %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  group_by(race) %>% 
  mutate(total_in_race = n()) %>% 
  mutate(percent_of_race = 1 / total_in_race) %>% 
  mutate(test_sum = sum(percent_of_race)) %>% 
  ggplot(aes(x = percent_of_race, y = race, fill = gender, color = gender)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Races and Gender", 
       x = "Percent", 
       y = "",
       fill = "Gender",
       color = "Gender")
```

# Countries

In the Airtable, multiple countries may be listed for one entry. These include 
places where people are involved or born, from what I gathered from a quick
reading of one source, usually Wikipedia. These countries are generally in order
of precedence. The first, and so primary, country for each entry is used here. 

An even greater  majority of entries are based in the U.S in this phase. Asian 
countries remain the most of those outside the U.S. With the exception of Australia,
there is at least one entry based in every continent.

```{r country-hist}
x %>% 
  mutate(country = str_replace_all(country, ",.*", "")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(country)))) +
  geom_bar() +
  labs(title = "Countries", 
       x = "Number",
       y = "")
```

# Parties

If a person has more than one party, each is counted here. For the most part, I 
listed parties for those who had one displayed on Wikipedia, i.e. primarily elected officials. 

The number of party members increased less than proportionally to the total number of entries, 
growing from 40 to 51, which may be due in part to the fact that I did not use 
Wikipedia as a source for all these entries. Democrats from a greater majority 
than they do last time. Several of the parties listed below are for foreign countries. 

```{r party-hist, warning = FALSE}
x %>% 
  separate_rows(party, sep = "\\s*,\\s*") %>% 
  mutate(party = str_replace_all(party, "Independent", "Independent (Irish)")) %>% 
  drop_na(party) %>% 
  ggplot(aes(y = party)) +
  geom_bar() +
  scale_y_discrete(limits = rev(c("D", "R", "I", "L", "Green", "Democratic Progressive Party",
                              "Independent (Irish)", "Kuomintang", "National Democratic Party of Tibet", 
                              "Popular Democratic", "Reform")),
                   labels = rev(c("Democrat", "Republican", "Independent", "Libretarian", 
                              "Green", "Democratic Progressive Party", "Independent (Irish)",
                              "Kuomintang", "National Democratic Party of Tibet", 
                              "Popular Democratic", "Reform"))) +
  labs(title = "Parties", 
       x = "Number", 
       y = "")
```

Men are the majority of politicians. Men now form a slightly greater majority 
among Democrats than in Phase 1. 

```{r party-gender}
x %>% 
  separate_rows(party, sep = "\\s*,\\s*") %>% 
  drop_na(party) %>% 
  mutate(party = str_replace_all(party, "Independent", "Independent (Irish)")) %>%
  filter(party %in% c("D", "R", "I", "L")) %>% 
  group_by(party) %>% 
  mutate(total_in_party = n()) %>% 
  mutate(percent_of_party = 1 / total_in_party) %>% 
  ggplot(aes(percent_of_party, party, fill = gender, color = gender)) +
  geom_col() +
  scale_y_discrete(labels = c("Democrat", "Independent", "Libretarian", "Republican")) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Parties and Gender",
       x = "Percent", 
       y = "", 
       fill = "Gender",
       color = "Gender")
```

Minorities now make up the majority of Democrats among the entries. 

```{r party-race}
x %>% 
   mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  filter(race %in% c("Jewish", "Black", "White", "Hispanic", "Asian")) %>% 
  mutate(race = factor(race,
                       levels = c("Black", "Asian", "Hispanic", "Jewish", "White"))) %>% 
  separate_rows(party, sep = "\\s*,\\s*") %>% 
  drop_na(party) %>% 
  mutate(party = str_replace_all(party, "Independent", "Independent (Irish)")) %>%
  filter(party %in% c("D", "R", "I", "L")) %>% 
  group_by(party) %>% 
  mutate(total_in_party = n()) %>% 
  mutate(percent_of_party = 1 / total_in_party) %>% 
  ggplot(aes(percent_of_party, party, fill = race, color = race)) +
  geom_col() +
  scale_y_discrete(labels = c("Democrat", "Independent", "Libretarian", "Republican")) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  scale_color_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  labs(title = "Parties and Race",
       x = "Percent", 
       y = "",
       fill = "Race",
       color = "Race")
```

# Roles

We made changes to how we defined roles in this Phase. We replaced the Politics 
category with Government, with includes both elected politicians and unelected officials,
e.g. cabinent members. The category also includes members of organizations such as
the U.N., though such also recieve the new role "International". The category 
still excludes judges, who remain in a separate "Judiciary" category.

"Public Interest Law" was rennamed to "Public Interest / Nonprofit Legal Services"
and now includes those who practice law for the public interest and members of 
think tanks and other NGOS.

Authors, musicians, journalists, and commentators have all been included into one 
new role, "Arts / Media / Entertainment." Media executives also recieve this role,
in addition to "Corporate/Business."

Among all the entries, government is the most frequently occuring role, slightly 
over 30% of all roles. Corporate/Business is now as popular a role as Judges, a big 
increase in the former and decrease in the latter compared to Phase 1. Law Firm
and Academia/Education remain smaller, but have increased more or less 
proportionally.


```{r role-hists, layout = "l-body-outset", fig.width= 8, fig.height = 5}
x %>% 
  separate_rows(category, sep = ",") %>% 
  ggplot(aes(y = fct_rev(fct_infreq(category)))) +
  geom_bar() +
  labs(title = "Roles of All Entries", 
       x = "Number", 
       y = "")
```

In most roles women outnumber men, most so in Academia / Education and International. 
Among Law Firm, Judiciary, and Activist, men are slightly more or as many.
Philanthropist is a small category which includes only men. 

```{r role-gender, layout = "l-body-outset", fig.width= 8, fig.height = 5}
x %>% 
  separate_rows(category, sep = ",") %>% 
  group_by(category) %>% 
  mutate(total_in_category = n()) %>% 
  filter(total_in_category > 1) %>% 
  mutate(percent_of_category = 1 / total_in_category) %>% 
  ggplot(aes(percent_of_category, category, fill = gender, color = gender)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(title = "Roles and Gender", 
       x = "Percent", 
       y = "", 
       fill = "Gender",
       color = "Gender")
```

```{r role-race, layout = "l-body-outset", fig.width= 8, fig.height = 5}
x %>% 
  mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
                          str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
                          str_detect(race, "Black") == TRUE ~ "Black", 
                          str_detect(race, "White") == TRUE ~ "White",
                          str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
                          str_detect(race, "Asian") == TRUE ~ "Asian",
                          str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>% 
  filter(race %in% c("Jewish", "Black", "White", "Hispanic", "Asian")) %>% 
  mutate(race = factor(race,
                       levels = c("Black", "Asian", "Hispanic", "Jewish", "White"))) %>% 
  separate_rows(category, sep = ",") %>% 
  group_by(category) %>% 
  mutate(total_in_category = n()) %>% 
  filter(total_in_category > 1) %>% 
  mutate(percent_of_category = 1 / total_in_category) %>% 
  ggplot(aes(percent_of_category, category, fill = race, color = race)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_fill_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  scale_color_manual(limits = c("Black", "Asian", "Hispanic", "Jewish", "White"),
                    values = c("black", "yellow2", "sienna", "steelblue3", "wheat")) +
  labs(title = "Roles and Race", 
       x = "Percent", 
       y = "", 
       fill = "Race", 
       color = "Race")

# x %>%
#   mutate(race = case_when(str_detect(race, "Jewish") == TRUE ~ "Jewish",
#                           str_detect(race, "Aboriginal") == TRUE ~ "Aboriginal",
#                           str_detect(race, "Black") == TRUE ~ "Black",
#                           str_detect(race, "White") == TRUE ~ "White",
#                           str_detect(race, "Hispanic") == TRUE ~ "Hispanic",
#                           str_detect(race, "Asian") == TRUE ~ "Asian",
#                           str_detect(race, "American Indian") == TRUE ~ "American Indian")) %>%
#   filter(race %in% c("Jewish", "Black", "White", "Hispanic", "Asian")) %>%
#   mutate(race = factor(race,
#                        levels = c("Black", "Asian", "Hispanic", "Jewish", "White"))) %>%
#   separate_rows(category, sep = ",") %>%
#   filter(category %in% c("Government", "Corporate / Business", "Judiciary",
#                          "Public Interest Law / Nonprofit Legal Services",
#                          "Academia / Education", "Law Firm")) %>%
#   group_by(race) %>%
#   mutate(total_in_race = n()) %>%
#   filter(total_in_race > 1) %>%
#   mutate(percent_of_race = 1 / total_in_race) %>%
#   ggplot(aes(percent_of_race, race, fill = category, color = category)) +
#   geom_col() +
#   scale_x_continuous(labels = scales::percent_format()) +
#   labs(title = "Race and Roles",
#        x = "Percent",
#        y = "",
#        fill = "Race",
#        color = "Race")
```

# Degrees

```{r degree-hist}
x %>% 
  mutate(degree = str_replace_all(degree, ",", " & ")) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(degree)))) +
  geom_bar() +
  labs(title = "Degrees", 
       x = "Number", 
       y = "")
```